---
title: "Time Series Analysis"
author: "Ramnath K (22BM6JP38), Vaibhav Goyal (22BM6JP54)"
date: "`r Sys.Date()`"
abstract: "This is abract for the project"
knit: (function(inputFile, encoding) {
      out_dir <- "Report";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
output: word_document
---

---
# Data is fetched from the Data subfolder.
# Knitted Output of the notebook will be stored in the Report subfolder.
# Flextable library is used to generate tables for word
---

***

# Initial Setup

#### Import Libraries:

1.  *astsa* - for Time Series Analysis
2.  *readxl* - for importing data from Excel file

```{r lib import}
library(astsa)
library(readxl)
```

```{r output setup, echo=FALSE}
# Flextable imported to create tables in output file
library(flextable)
# Table caption set to bottom
knitr::opts_chunk$set(tab.topcaption = FALSE)
# Avoid scientific notations in the plots
options(scipen = 999)
```


#### Import Data
```{r data import}
data <- read_xlsx("./Data/US_MRTR.xlsx", sheet = 'Sales')
```
```{r table code, echo=FALSE}
ft <- 
  flextable(head(data)) |>
  set_table_properties(layout = "autofit") |>
  set_caption(caption = "Sample from data") |>
  colformat_num(j = 'Year', big.mark = "")
ft
```

Before proceeding, let us perform some minor transformations:

1.  Change name of the sales column for easier reference.
2.  Change unit of the sales from Millions to Billions.

```{r data tansform}
colnames(data)[3] <- 'Sales'
data$Sales <- data$Sales / 1000 # To change scale from $ Millions to $ Billions
```

```{r table_2 code, echo=FALSE}
ft <- 
  flextable(head(data)) |>
  set_table_properties(layout = "autofit") |>
  set_caption(caption = "Sample from data after transformations") |>
  colformat_num(j = 'Year', big.mark = "")
ft
```
# Analysis
#### Define time series
The *ts()* function from the *astsa* helps us define an indexed time series from out sales data.\
Sometimes when plotting the data, it is helpful to have the actual time intervals on the on the axis, for this purpose we also create a time series indexed by the time intervals *(sales.ts in the below code)*.
```{r ts}
sales <- ts(data = data$Sales)
sales.ts <- ts(data = data$Sales, frequency = 12, start = c(1992,1))
```
#### Plot time series
```{r plot, echo=FALSE, fig.width=9, fig.height=6}
plot(sales.ts, type="o", pch = 20, ylab = "Sales (in $ 'B)", main = "Monthly New Car Sales in US [Jan 1992 - Oct 2022]")
rect(xleft=2008,xright = 2010, ybottom = par("usr")[3], ytop=par("usr")[4], density=10, col = "red")
rect(xleft=2020,xright = 2021, ybottom = par("usr")[3], ytop=par("usr")[4], density=10, col = "blue")

```

On plotting the time series data, we observe a clear increasing trend and a seasonality in the data.

We observe a departure from the general trend for the years 2008 to 2010 which may be attributed to the global economic crisis of 2008. *(Highlighted in red region in the above plot)*

We also observe a sharp drop in the sales in the years 2020 to 2021 which can be attributed to the CoVID-19 pandemic. *(Highlighted in blue region in the above plot)*

For our initial initial analysis, we will ignore these two anomalies and proceed with the data as is.

#### AR(1) Model
Since there is upward trend, we first need to make it stationary to fit an AR(1) model. To create a (possibly) stationary series, we’ll examine the first differences $y_t = x_t - x_{t-1}$.
```{r}
lag1.plot(sales,1)
```
```{r, echo=FALSE, fig.width=9, fig.height=6}
acf(sales)
```

```{r, echo=FALSE, fig.width=9, fig.height=6}
pacf(sales, main = 'New Car Sales')
```
