---
title: "Time Series Analysis"
author: "Ramnath K (22BM6JP99), Vaibhav Goyal (22BM6JP54)"
date: "`r Sys.Date()`"
abstract: "This is abract for the project"
knit: (function(inputFile, encoding) {
      out_dir <- "Report";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
output:
  word_document: default
---

---
# Data is fetched from the Data subfolder.
# Knitted Output of the notebook will be stored in the Report subfolder.
# Flextable library is used to generate tables for word

# ```{r}
# library(flextable)
# ```

---

***

# Initial Setup

### Import Libraries:

1.  *astsa* - for Time Series Analysis
    i.
2.  *readxl* - for importing data from Excel file
```{r lib import}
library(astsa)
library(readxl)
```

### Import Data

```{r data import}
data <- read_xlsx("./Data/BSE_Small_Cap.xlsx", sheet = 'Data')
```


```{r datainclude=FALSE}
flextable::flextable(head(data, 10))
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot